---
title: "Using Remote Sensing Data to Understand Fire Ignition Risk in Victoria"
author: "Helen Evangelista, Brenwin Ang, Di Cook, Emily Dodwell, Chris Volinsky, Weihao Li"
date: "ACEMS Retreat, Nov 3 2021"
pdflink: ""
bgimg: "images/Werombi_Bushfire.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/monash-logo.css"
      - "assets/monash-brand.css"
      - "assets/monash-fonts.css"
      - "assets/styles.css" # small improvements
      - "assets/custom.css" # add your own CSS here!
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    lib_dir: libs
    includes:
      in_header: "assets/custom.html"
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: github 
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
Sys.setlocale("LC_ALL","English")
```

```{r titleslide, child="components/titleslide.Rmd"}
```

---
# IDEAS Challenge


*Motivation*: The collaboration would develop risk models and visualisations, to monitor potential fire ignitions and track fires from satellite hotspots, in real-time, for the 2021-2022 Victorian bushfire season. These will be made these publicly available by extending the existing web app.

*Team*:


---

# 🗺 Background

1. Algorithm to detect bushfire ignition from hotspot data, published in R package `spotoroo` and upcoming R Journal article
2. Model to predict the cause of bushfire ignition, using historical data, and metadata, temperature, precipitation, winds, radiation, vegetation/fuel layer, roads, recreation sites, CFA stations. 
3. Prediction of the causes of the 2019-2020 Australia bushfires
4. A complete and adaptable workflow for monitoring and understanding new ignitions from hotspot data
5. Shiny app for exploration of historical fire origins, predicted causes of 2019-2020 fires and future fire risk maps
6. All work conducted with open data and open source software.

---
# Shiny app: https://ebsmonash.shinyapps.io/VICfire/

```{r out.width="110%"}
knitr::include_app("https://ebsmonash.shinyapps.io/VICfire/", height = "550px") 
```

---
# Data sources update

- 📡 Remote sensing data for 2016/2017-2020/2021 seasons converted into fire ignition locations: Japan Aerospace Exploration Agency  provides a hotspot product (reflected energy from the earth) taken from the **Himawari-8** satellite.
- 🔥 Historical fire origins for 2019/2020: can compare our predictions for that season now.
- 🌦️🌿⛺ Metadata: updated and expanded for all seasons


```{r eval=FALSE}
if (!file.exists("images/hotspots_before.png")){
  library(rnaturalearth)
  library(ggthemes)
  library(lubridate)
  au_map <- ne_states(country = 'Australia', returnclass = 'sf')
  vic_map <- au_map[7,]
  VIC_hotspots_before_clustering <- read_csv("data/VIC_hotspots_before_clustering.csv") %>%
    mutate(date = as.Date("2019-10-01") + hours(time_id)) %>%
    mutate(month = month(date, label = TRUE)) %>%
    mutate(month = factor(month, levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar")))

  p <- ggplot(VIC_hotspots_before_clustering) +
    geom_sf(data = vic_map, fill = "grey99", 
            colour = "black", size=0.1) +
    geom_point(aes(lon, lat), size = 1, 
               alpha=0.3, colour = "#fd8d3c") +
    facet_wrap(~month, ncol=3) +
    theme_map() + 
    theme(strip.background = element_rect(fill="white"),
          panel.border = element_rect(colour="black", fill=NA))
  
  ggsave(p, filename = "images/hotspots_before.png", height = 3.5, width = 6, dpi = 300)
}
```


```{r eval=FALSE}
if (!file.exists("images/hotspots_after.png")){
  library(rnaturalearth)
  library(ggthemes)
  library(lubridate)
  au_map <- ne_states(country = 'Australia', returnclass = 'sf')
  vic_map <- au_map[7,]
  VIC_hotspots_before_clustering <- read_csv("data/VIC_hotspots_before_clustering.csv") %>%
    mutate(date = as.Date("2019-10-01") + hours(time_id)) %>%
    mutate(month = month(date, label = TRUE)) %>%
    mutate(month = factor(month, levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar")))

  p <- ggplot(VIC_hotspots_before_clustering) +
    geom_sf(data = vic_map, fill = "grey99", 
            colour = "black", size=0.1) +
    geom_point(aes(lon, lat), size = 2, 
               alpha=0.05, colour = "grey90") +
    facet_wrap(~month, ncol=3) +
    theme_map() + 
    theme(strip.background = element_rect(fill="white"),
          panel.border = element_rect(colour="black", fill=NA))
  
  VIC_hotspots_after_clustering <- read_csv("data/VIC_hotspots_after_clustering.csv")
  hotspots <- read_csv("data/VIC_hotspots_raw.csv")
  
  VIC_hotspots_before_clustering$fire_id <- VIC_hotspots_after_clustering$fire_id
  VIC_hotspots_before_clustering$date <- hotspots$`#obstime`
  
  VIC_hotspots_after_clustering <- VIC_hotspots_before_clustering %>%
    group_by(fire_id) %>%
    summarise(time_id = min(time_id)) %>%
    left_join(VIC_hotspots_before_clustering) %>%
    group_by(fire_id, time_id) %>%
    summarise(lon = mean(lon), lat = mean(lat), 
              date = mean(date)) %>%
    mutate(month = month(date, label = TRUE)) %>%
    mutate(month = factor(month, levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar")))
  
  p <- p +
    geom_point(data = VIC_hotspots_after_clustering, 
               aes(lon, lat), alpha = 1, 
               size = 0.2, colour="#fd8d3c") 
  
  ggsave(p, filename = "images/hotspots_after.png", height = 3.5, width = 6, dpi = 300) 
}

```


```{r eval=FALSE}
if (!file.exists("images/ignition_summary.png")){
library(tidyverse)
library(lubridate)
library(ggthemes)
library(rnaturalearth)

set.seed(357)

# Read in training data
training <- read_csv("data/training.csv")

training <- training %>%
  filter(!CAUSE %in% c("BURNING BUILDING",
                       "WASTE DISPOSAL, INDUSTRIAL, SAWMILL, TIP",
                       "WASTE DISPOSAL, DOMESTIC",
                       "BURNING VEHICLE, MACHINE",
                       "BURNING BUILDING")) %>%
  filter(new_cause != "other") %>%
  filter(new_cause != "relight")


training <- select(training, -c(EVENTID:FIRE_NUM), -id, -CAUSE, -FOREST, -FOR_CODE, -FOR_CAT)

training <- mutate(training,
                   year = factor(year(FIRE_START)),
                   month = factor(month(FIRE_START), levels = c(10,11,12,1,2,3)),
                   day = factor(day(FIRE_START), levels = c(1:31)),
                   wod = factor(wday(FIRE_START), levels = c(1:7)))

training <- filter(training, month %in% c(10,11,12,1,2,3))


training <- na.omit(training)

training <- mutate(training, new_cause = ifelse(new_cause == "accidental_human", "accident", new_cause)) %>%
  mutate(new_cause = ifelse(new_cause == "burning_off_human", "burning_off", new_cause)) %>%
  mutate(new_cause = factor(new_cause)) %>%
  mutate(FOR_TYPE = factor(FOR_TYPE))

training <- na.omit(training)

training <- mutate(training,
                   log_dist_cfa = log(dist_cfa),
                   log_dist_camp = log(dist_camp),
                   log_dist_road = log(dist_road),
                   COVER = factor(COVER),
                   HEIGHT = factor(HEIGHT))

training <- rename(training, cause = new_cause)
training <- mutate(training,
                   cause = fct_relevel(cause,
                                       "lightning",
                                       "accident",
                                       "arson",
                                       "burning_off"))

training <- na.omit(training)

p <- training %>%
        mutate(cause = factor(tools::toTitleCase(as.character(cause)), 
     levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
    ggplot() +
     geom_histogram(aes(cause, fill = cause), stat = "count") +
    theme_minimal(base_size = 40) +
    ggtitle("Histogram of ignitions by causes") +
    theme(legend.position = "none") +
    scale_fill_brewer(palette = "RdBu") +
    xlab("Cause") +
    ylab("Count") 
  
  ggsave(p, filename = "images/ignition_summary.png", height = 10, width = 12, dpi = 600)
  
p <- training %>%
  mutate(cause = factor(tools::toTitleCase(as.character(cause)), levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
    ggplot() +
  geom_bar(data = select(training, -cause), aes(as.Date(paste0(as.character(year), "-01-01")))) +
    geom_bar(aes(as.Date(paste0(as.character(year), "-01-01")), fill = cause)) +
    facet_wrap(~cause) +
    theme_minimal(base_size = 40) +
    ggtitle("Histogram of ignitions by year") +
    theme(legend.position = "none") +
    scale_fill_brewer(palette = "RdBu") +
    xlab("Year") +
    ylab("Count") +
    scale_x_date(date_labels = "%y")
  
  ggsave(p, filename = "images/ignition_year.png", height = 10, width = 12, dpi = 600)
}
  
```



```{r eval=FALSE}
if (!file.exists("images/density.png")){
  library(tidyverse)
library(lubridate)
library(ggthemes)
library(rnaturalearth)
au_map <- ne_states(country = 'Australia', returnclass = 'sf')
vic_map <- au_map[7,]

set.seed(357)

# Read in training data
training <- read_csv("data/training.csv")

training <- training %>%
  filter(!CAUSE %in% c("BURNING BUILDING",
                       "WASTE DISPOSAL, INDUSTRIAL, SAWMILL, TIP",
                       "WASTE DISPOSAL, DOMESTIC",
                       "BURNING VEHICLE, MACHINE",
                       "BURNING BUILDING")) %>%
  filter(new_cause != "other") %>%
  filter(new_cause != "relight")


training <- select(training, -c(EVENTID:FIRE_NUM), -id, -CAUSE, -FOREST, -FOR_CODE, -FOR_CAT)

training <- mutate(training,
                   year = factor(year(FIRE_START)),
                   month = factor(month(FIRE_START), levels = c(10,11,12,1,2,3)),
                   day = factor(day(FIRE_START), levels = c(1:31)),
                   wod = factor(wday(FIRE_START), levels = c(1:7)))

training <- filter(training, month %in% c(10,11,12,1,2,3))


training <- na.omit(training)

training <- mutate(training, new_cause = ifelse(new_cause == "accidental_human", "accident", new_cause)) %>%
  mutate(new_cause = ifelse(new_cause == "burning_off_human", "burning_off", new_cause)) %>%
  mutate(new_cause = factor(new_cause)) %>%
  mutate(FOR_TYPE = factor(FOR_TYPE))

training <- na.omit(training)

training <- mutate(training,
                   log_dist_cfa = log(dist_cfa),
                   log_dist_camp = log(dist_camp),
                   log_dist_road = log(dist_road),
                   COVER = factor(COVER),
                   HEIGHT = factor(HEIGHT))

training <- rename(training, cause = new_cause)
training <- mutate(training,
                   cause = fct_relevel(cause,
                                       "lightning",
                                       "accident",
                                       "arson",
                                       "burning_off"))

training <- na.omit(training)

  training %>%
  mutate(cause = factor(tools::toTitleCase(as.character(cause)), levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
  ggplot() +
      geom_density_2d_filled(aes(lon, lat), contour_var = "ndensity") +
      geom_sf(data = vic_map, fill = NA, col = "white") +
      facet_wrap(~cause)+
      ggtitle("2D conditional density plot of historical bushfire ignitions") +
    theme_map(base_size = 20) +
    theme(legend.position = "right",
          plot.margin = unit(c(0, 0, 0, 0), "pt")) +
    guides(fill = guide_legend(reverse = T)) -> p
  
  ggsave(p, filename = "images/density.png", height = 6, width = 12, dpi = 1200)
}


    
```


```{r eval=FALSE}
if (!file.exists("images/density_cfa.png")){
    library(tidyverse)
library(lubridate)
library(ggthemes)
library(rnaturalearth)
au_map <- ne_states(country = 'Australia', returnclass = 'sf')
vic_map <- au_map[7,]

set.seed(357)

# Read in training data
training <- read_csv("data/training.csv")

training <- training %>%
  filter(!CAUSE %in% c("BURNING BUILDING",
                       "WASTE DISPOSAL, INDUSTRIAL, SAWMILL, TIP",
                       "WASTE DISPOSAL, DOMESTIC",
                       "BURNING VEHICLE, MACHINE",
                       "BURNING BUILDING")) %>%
  filter(new_cause != "other") %>%
  filter(new_cause != "relight")


training <- select(training, -c(EVENTID:FIRE_NUM), -id, -CAUSE, -FOREST, -FOR_CODE, -FOR_CAT)

training <- mutate(training,
                   year = factor(year(FIRE_START)),
                   month = factor(month(FIRE_START), levels = c(10,11,12,1,2,3)),
                   day = factor(day(FIRE_START), levels = c(1:31)),
                   wod = factor(wday(FIRE_START), levels = c(1:7)))

training <- filter(training, month %in% c(10,11,12,1,2,3))


training <- na.omit(training)

training <- mutate(training, new_cause = ifelse(new_cause == "accidental_human", "accident", new_cause)) %>%
  mutate(new_cause = ifelse(new_cause == "burning_off_human", "burning_off", new_cause)) %>%
  mutate(new_cause = factor(new_cause)) %>%
  mutate(FOR_TYPE = factor(FOR_TYPE))

training <- na.omit(training)

training <- mutate(training,
                   log_dist_cfa = log(dist_cfa),
                   log_dist_camp = log(dist_camp),
                   log_dist_road = log(dist_road),
                   COVER = factor(COVER),
                   HEIGHT = factor(HEIGHT))

training <- rename(training, cause = new_cause)
training <- mutate(training,
                   cause = fct_relevel(cause,
                                       "lightning",
                                       "accident",
                                       "arson",
                                       "burning_off"))

training <- na.omit(training)

p <-training %>%
  mutate(cause = factor(tools::toTitleCase(as.character(cause)), levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
    gather(key = "metric", value = "distance", log_dist_cfa, log_dist_road) %>%
    mutate(metric = ifelse(metric == "log_dist_cfa", "Log distance to the nearest CFA station", "Log distance to the nearest road")) %>%
    ggplot() +
    geom_density(aes(distance, colour = cause, 
                     fill = cause), size = 1.5,
                 alpha=0.3) +
    theme_minimal(base_size = 20) +
    theme(legend.position = "bottom") +
    xlab("Log distance") +
    scale_color_brewer(palette = "RdBu") +
    scale_fill_brewer(palette = "RdBu") +
    facet_wrap(~metric, scales = "free_x") 
  
  ggsave(p, filename = "images/density_cfa.png", height = 6, width = 12, dpi = 600)
  
}
```


---
# 📐 Modelling risk

.monash-blue[ Something]



```{r eval=FALSE}
library(kableExtra)
data.frame(model = c("Multinomial logistic regression", "GAM multinomial logistic regression", "Random forest", "XGBoost"), 
           Accuracy = c(0.53, 0.68, 0.75, 0.74),
           AUC = c(0.74, 0.82, 0.88, 0.88)) %>%
  `colnames<-`(c("Model", "Accuracy", "Muti-class AUC")) %>%
  knitr::kable("html")  %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(font_size = 20)
```



```{r eval=FALSE}
data.frame(Lightning = c(703, 51, 18, 5, 777), 
           Accident = c(77, 494, 55, 8, 634),
           Arson = c(50, 89, 175, 11, 325),
           Burning_off = c(44, 38, 22, 32, 136),
           Total = c(874, 672, 270, 56, 1872)) %>%
  mutate(Lightning = c(paste0(Lightning[1:4], " (", round(Lightning[1:4]/Lightning[5], 2), ")" ), paste0(Lightning[5]))) %>%
  mutate(Accident = c(paste0(Accident[1:4], " (", round(Accident[1:4]/Accident[5], 2), ")" ), paste0(Accident[5]))) %>%
  mutate(Arson = c(paste0(Arson[1:4], " (", round(Arson[1:4]/Arson[5], 2), ")" ), paste0(Arson[5]))) %>%
  mutate(Burning_off = c(paste0(Burning_off[1:4], " (", round(Burning_off[1:4]/Burning_off[5], 2), ")" ), paste0(Burning_off[5]))) %>%
  `row.names<-`(c("Prediction:Lightning", "Prediction:Accident", "Prediction:Arson", "Prediction:Burning_off", "Total")) %>%
  knitr::kable("html")  %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(font_size = 20)
  
```


```{r eval=FALSE}
if (!file.exists("images/varimp.png")){
  
   explaination <- readRDS("data/lime_ex2.rds")
  
  name_dict <- list("log_dist_camp" = "Log distance to the nearest camping site",
       "log_dist_cfa" = "Log distance to the nearest CFA station",
       "log_dist_road" = "Log distance to the nearest road",
       "lon" = "longitude",
       "lat" = "Latitude",
       "ase180" = "Average solar exposure in the past 180 days",
       "arf360" = "Average rainfall in the past 360 days",
       "ase90" = "Average solar exposure in the past 90 days",
       "aws_m24" = "Average wind speed in the past 24 months",
       "aws_m12" = "Average wind speed in the past 12 months")
  
  
  explaination %>%
  mutate(label = factor(tools::toTitleCase(label), levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
    mutate(feature = unlist(name_dict[feature])) %>%
  ggplot() +
    geom_boxplot(aes(feature_weight, feature), outlier.size = 0.5) +
    geom_vline(xintercept = 0, col = "red") +
    ylab("feature") +
    facet_wrap(~label, ncol = 2) +
    xlab("Feature Weight") +
    ylab("Feature") +
    theme_minimal(base_size = 20) -> p
  ggsave(p, filename = "images/varimp.png", height = 6, width = 12, dpi = 600)
}
```


```{r out.width="100%", eval=FALSE}
library(ggthemes)
library(rnaturalearth)
library(lubridate)
au_map <- ne_states(country = 'Australia', returnclass = 'sf')
vic_map <- au_map[7,]

read_csv("data/prediction_2019-2020.csv") %>%
  mutate(cause = factor(tools::toTitleCase(cause), levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
  ggplot() +
  geom_sf(data = vic_map, fill = "grey99") +
  geom_point(aes(lon, lat, col = cause, text = paste0("Mon: ", month.abb[month(time)])), size = 1.5, alpha = 1) +
  theme_map(base_size = 20) +
  theme(legend.position = "right") +
  # facet_wrap(~factor(month.abb[month(time)], levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar"))) +
  scale_color_brewer(palette = "RdBu")-> p

library(plotly)


ggplotly(p, tooltip = c("colour", "text"))

```

---
# 🔭 Plan for completion


```{r eval=FALSE}
read_csv("data/prediction_2019-2020.csv") %>%
  mutate(cause = factor(tools::toTitleCase(cause), levels = c("Lightning", "Accident", "Arson", "Burning_off"))) %>%
  mutate(month = factor(month.abb[month(time)], levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar"))) %>%
  group_by(cause, month) %>%
  count() %>%
  ungroup() %>%
  rename(Cause = cause) %>%
  spread(month, n) -> temp

temp[is.na(temp)] <- 0

temp$Total <- temp$Oct + temp$Nov + temp$Dec + temp$Jan + temp$Feb + temp$Mar
temp$Total <- paste0(temp$Total, " (", round(temp$Total/sum(temp$Total), 2), ")")

temp  %>%
  knitr::kable("html")  %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(font_size = 20)
  
```

---

# Relevance to Telecommunications Industry

- Extreme weather and climate-related events pose significant risks to cellular network infrastructure, e.g.:
	+ "Wildfire threat: Bay Area cell phone, internet service could go out, too" (*San Francisco Chronicle* - August 20, 2020)
	+ "911 calls after Ida went unanswered in New Orleans due to ‘antiquated’ technology" (*The Washington Post* - August 30, 2021)
- Necessary for telecom providers to address impact of climate perils including wind, flood, and fire on customer connectivity

---

# Benefits of Collaboration to AT&T

- As the severity of climate disasters are expected to increase, AT&T is committed to staying ahead of these dangers:
	+ Developed *Climate Change Analysis Tool* with the US Department of Energy's Argonne National Laboratory
	+ Published a white paper entitled *The Road to Climate Resiliency*
* ACEMS research collaboration to predict wildfire risk is aligned with AT&T's climate resiliency efforts
* Development of tools and methods for analytics and visualization of wildfire incidence are generalizable and extendable to other geographies



---
```{r endslide, child="components/endslide.Rmd"}
```


