---
title: "ida"
author: "Brenwin"
date: "19/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# --- load library
library(tidyverse)
```

# read CFA data 

    data source:
- https://discover.data.vic.gov.au/dataset/low-bushfire-rating-areas1

    about data:
- define areas of low fire hazard ratings
- areas NOT covered by this layer; are rated as having high fire hazard rating (for purposes of Electiricty Safety Act 1998)
    
- CFA; assign low/high fire ratings; to land areas
  - required by law (Electricity Safety Act 1998)
  - used in conjunction with Electricity Safety (Bushfire Mitigation) Regulations 2013; to prescribe low/high bushfire risk areas 


    low/high fire hazard rating areas
Low fire hazard rating area
- usually restricted to irrigated areas & towns; where block sizes are small & well maintained
  - e.g. high density residential areas// industrial & commercial urban areas// high moisture crops (e.g. market gardens)// irrrigated farmland// vineyards// golf courses// sporting ovas
  
High fire hazard rating area
- areas; not covered by low hazard rating areas
- areas; said to contain sufficient fuel on ground; to carry a fire

Hazardous bushfire risk area (HBRA): area assigned a fire hazard rating of "high"
Low bushfire risk area (LBRA): areas assigned a fire hazard rating of "low"
  

```{r}
lbra <- sf::st_read(here::here("data/LBRA/Low_bushfire_rating_areas.shp"))

sf::st_geometry(lbra) %>% 
  plot()

tmap::tm_shape(lbra) +
  tmap::tm_polygons()
```

# grassland curing

    data source
- https://portal.tern.org.au/grassland-curing-modis-states-coverage/21005

    About grassland curing
- green lush grassland; unlikely to burn // dry brown grassland; likely; burn
- grassland curing: natural process; change grassland; from green(0%) to brown(100%)
  - 10-30%: unlikely to burn
  - 40-60%: in between
  - 70-100%: likely; burn rapidly 

```{r}
raster::raster(here::here("data/grassland_curing.tif")) %>% 
  raster::getValues()
```







