---
title: "eda"
author: "Brenwin"
date: "06/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# --- load libraries
library(tidyverse)
library(ozmaps)
library(tmap)
library(raster)
library(sf)
library(sp)
library(spatstat)
```

# read in data
```{r, eval=FALSE}
# === read in data sets
load("save.RData")
```

## historical data (`training`) *historical bushfire (less reliable)
```{r}
training <- read_csv("data/training.csv")

training <- training %>%
  filter(!CAUSE %in% c("BURNING BUILDING",
                       "WASTE DISPOSAL, INDUSTRIAL, SAWMILL, TIP",
                       "WASTE DISPOSAL, DOMESTIC",
                       "BURNING VEHICLE, MACHINE",
                       "BURNING BUILDING")) %>%
  filter(new_cause != "other") %>%
  filter(new_cause != "relight")


training <- select(training, -c(EVENTID:FIRE_NUM), -id, -CAUSE, -FOREST, -FOR_CODE, -FOR_CAT)

training <- training %>% 
  mutate(year = factor(lubridate::year(FIRE_START)),
         month = factor(lubridate::month(FIRE_START)),
                        # levels = c(10,11,12,1,2,3),
         day = factor(lubridate::day(FIRE_START), 
                      levels = c(1:31)),
         wod = factor(lubridate::wday(FIRE_START), 
                      levels = c(1:7)))

# training <- training %>% 
#   filter(month %in% c(10,11,12,1,2,3))

training <- na.omit(training)

training <- training %>% 
  mutate(new_cause = ifelse(test = new_cause == "accidental_human",
                            yes = "accident",
                            no = new_cause)) %>%
  mutate(new_cause = ifelse(test = new_cause == "burning_off_human", 
                            yes = "burning_off", 
                            no = new_cause)) %>%
  mutate(new_cause = factor(new_cause)) %>%
  mutate(FOR_TYPE = factor(FOR_TYPE))

training <- na.omit(training)

training <- training %>% 
  mutate(log_dist_cfa = log(dist_cfa),
         log_dist_camp = log(dist_camp),
         log_dist_road = log(dist_road),
         COVER = factor(COVER),
         HEIGHT = factor(HEIGHT))

training <- training %>% 
  rename(cause = new_cause)

training <- mutate(training,
                   cause = forcats::fct_relevel(cause,
                                                "lightning",
                                                "accident",
                                                "arson",
                                                "burning_off"))

training <- na.omit(training)

training <- training %>% 
  select(-dist_road, -dist_cfa, -dist_camp, -FIRE_START)
```

## import random forest model (`final_model`)
```{r}
final_model <- readr::read_rds("data/Final_model.rds")
```

## training data `predict_x` *satellite data
```{r}
# Read in predict data
predict_x <- readr::read_csv("data/predict_x.csv")

predict_x <- predict_x %>% 
  mutate(log_dist_cfa = log(dist_cfa),
         log_dist_camp = log(dist_camp),
         log_dist_road = log(dist_road))
```

## use RF model; predict bushfire ignitions cause; in satellite data (`predict_x`) and historical data (`training`)
```{r}
# === use RF model; predict bushfire `cause`

# --- predict cause; in training *historical data 
training_pred <- training %>% 
  # use RF model; predict bushfire cause 
  mutate(cause = predict(final_model,
                         newdata = training))

# --- predict cause; in `predict_x` *satellite data
predict_x <- predict_x %>% 
  # use RF model; predict bushfire cause 
  mutate(cause = predict(final_model,
                         newdata = predict_x),
         # incldue `month` and `year` variable
         month = factor(lubridate::month(time)),
         year = factor(lubridate::year(time)))
```


```{r}
# e.g. === plot results
au_map <- rnaturalearth::ne_states(country = 'Australia', 
                                   returnclass = 'sf')

vic_map <- au_map[7, ]

# === plot maps

# `training_pred` *historical data 
ggplot() +
  geom_sf(data = vic_map) +
  geom_point(aes(x = lon,
                 y = lat,
                 colour = cause),
             alpha = 0.8,
             data = training_pred) +
  facet_wrap(~ month,
             ncol = 3) +
  ggthemes::theme_map() +
  theme(legend.position = "right")

# `predict_x` *satellite data
ggplot() +
  geom_sf(data = vic_map) +
  geom_point(aes(x = lon,
                 y = lat,
                 colour = cause),
             alpha = 0.8,
             data = predict_x) +
  facet_wrap(~ month,
             ncol = 3) +
  # themes
  ggthemes::theme_map() +
  theme(legend.position = "right")
```

## `cause` by `year` bar plot (temporal)
```{r}
# ---------- `predict_x` *satellite data

# --- `cause` by `year`
predict_x %>% 
  group_by(year) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(year),
               y = total)) +
  geom_col(aes(x = factor(year),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw()

# --- `cause` by `month`
predict_x %>% 
  group_by(month) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(month),
               y = total)) +
  geom_col(aes(x = factor(month),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw()
```

```{r}
# ---------- `training_pred` *historical data
training_pred %>% 
  group_by(year) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(year),
               y = total)) +
  geom_col(aes(x = factor(year),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw()

# --- `cause` by `month`
training_pred %>% 
  group_by(month) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(month),
               y = total)) +
  geom_col(aes(x = factor(month),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw()
```



# ignition grid 

Goal
> form spatially weighted grid cells; with each grid cell
  - given ID 
  - based on historical data & 2019-2020 satellite data; compute 
    - fire ignition counts
    - proportion $\frac{\# fire~ ignitions}{\# total}$

> examine the grid cells
  - e.g. is historical data understating bushfire in remote areas?
  - make side-by-side dot plots of proportions; by cell id
  

## create Victoria map `sf` (grid's outline)

> Victoria map; forms spatial grid's outline 

```{r}
# === Victoria map (sfdf MULTIPOLYGON)
vic_map_sf <- ozmaps::ozmap_states %>% 
  filter(NAME == "Victoria")

# --- project crs
vic_map_sf <- sf::st_transform(vic_map_sf,
                               crs = 4326)

# === create cause_sf (sfdf POINT); from predictions *satellite data
cause_sf <- predict_x %>% 
  select(lon, lat, cause, year, FOR_CAT) %>%  # extract year; from date
  sf::st_as_sf(coords = c("lon", "lat"),
               crs = 4326)

# === create cause_training_sf (sfdf POINT); *historical (`training`) data
cause_training_sf <- training_pred %>% 
  select(lon, lat, cause, year) %>% 
  sf::st_as_sf(coords = c("lon", "lat"),
               crs = 4326)
```

## create spatial grid `RasterBrick`


### create `RasterBrick`; cropped to Victoria map window

> `RasterBrick`; designed; store multiple raster (with same extents(bbox) & dimensions)
  - i.e. multi-layer raster

```{r}
# === vic_raster 

# --- create `vic_raster` object *to be toggled
vic_raster <- raster::brick(
  # no. of rows & columns (directly linked to resolution of grid cell)
  nrows = 20,
  ncols = 20,
  
  # bbox (bounding box of Victoria)
  xmn = 140.9617,
  xmx = 149.9763,
  ymn = -39.13396,
  ymx = -33.99605,
  
  # crs
  crs = 4326,
  
  # set `raster` values (rowwise)
  # vals = seq(from = 1, to = 400000, by = 1000) 
  )

# --- mask raster; to only Victorian map 

# change vic_map_sf to `sp` object 
# *`raster` package; NOT compatible with `sf` yet; so; need; change to `sp` 
vic_map_sp <- as(vic_map_sf, 
                 Class = "Spatial")

# mask (*think: crop to polygon shape) raster; to only Victorian map (`vic_map_sp`)
vic_raster_crop <- vic_raster %>% 
  raster::mask(mask = vic_map_sp)
```

### ignition Rules
> to be set; when more data and information on appropriate rules are determined
   - e.g. filter to seasons 
    - fire; more prone in summer (increase bushfire risk)
   - e.g. categorise each grid cell; according to fuel(vegetation)
    - e.g. fire cannot occur on water (no bushfire risk)

```{r}
ignition_rules_raster <- raster::brick(
  # no. of rows & columns (directly linked to resolution of grid cell)
  nrows = 20,
  ncols = 20,
  
  # bbox (bounding box of Victoria)
  xmn = 140.9617,
  xmx = 149.9763,
  ymn = -39.13396,
  ymx = -33.99605,
  
  # crs
  crs = 4326,
  
  # set `raster` values (rowwise)
  # vals = seq(from = 1, to = 400000, by = 1000) 
  )

# --- convert `sp` to `sf` Polygons`; to conduct spatial join 
ignition_rules_sp <- as(ignition_rules_raster, "SpatialPolygonsDataFrame")
ignition_rules_sf <- sf::st_as_sf(ignition_rules_sp) 

ignition_rules_sf_values <- sf::st_join(ignition_rules_sf, cause_sf) 
```

```{r}
# --- plot ignition rule 
tmap::tm_shape(ignition_rules_sf_values) +
  tm_polygons(col = "FOR_CAT") +
  tm_shape(vic_map_sf) +
  tm_borders(lwd = 3)
```


### Set values to each grid cell

> set values to ignition grid cell
  - `fire_count`: count the number of fire ignitions; per grid 
  - `fire_prop`: # ignitions in grid cell / # total no. of ignitions

* note: good to separate satellite data (`predict_x`) with historical data (`training_pred`)
  - so; able; parse them out; when examining grid 

```{r}
# ========== convert spdf to SpatialPoints; to be able; use `rasterize` (count; no. of pts.; in each raster cell)

# ---------- predict_x *satellite data
cause_spdf <- as(cause_sf, Class = "Spatial")
cause_sp <- as(cause_spdf, "SpatialPoints")

# ---------- training_pred *historical data

# will filter `sf` object; before turning int `spdf` object in app  
# e.g. filter to year 2000 only (user-specification!)
cause_training_sf <- cause_training_sf %>% 
  filter(year == 2000)

cause_training_spdf <- as(cause_training_sf, Class = "Spatial")
cause_training_sp <- as(cause_training_spdf, "SpatialPoints")
```

> NOTE: BEFORE creating `spdf` object; is where; filter to `year` etc. (user specifications)


```{r}
# --- count the number of points in each raster cell; via `raster::rasterize`

# predict_x *satellite data 
ignition_rasterize <- raster::rasterize(x = cause_sp,# `SpatialPoints` object
                                        y = vic_raster_crop, # `Raster` object 
                                        fun = "count") 


ignition_rasterize_training <- raster::rasterize(x = cause_training_sp,# `SpatialPoints` object
                                                 y = vic_raster_crop, # `Raster` object 
                                                 fun = "count") 

# ----- compute; in each grid cell
# • `fire_count`: no. of points
# • `fire_prop`: no. of points / total points (bushfire ignitions)

# predict_x *satellite data 
ignition_raster_values <- raster::getValues(ignition_rasterize) %>% # extract count values
  tibble(fire_count = .) %>% 
  replace_na(list(fire_count = 0)) %>% # replace NA with 0 (no fire ignitions) 
  mutate(id = 1:nrow(.),
         .before = fire_count) %>% 
  mutate(fire_prop = fire_count / sum(fire_count)) %>% # no. of fire ignitions in a cell / total no. of fire ignitions
  as.matrix() # turn intro matrix (each column = 1 layer in `RasterBrick`)

# `training` *historical data 
ignition_raster_values_training <- raster::getValues(ignition_rasterize_training) %>% # extract count values
  tibble(fire_count = .) %>% 
  replace_na(list(fire_count = 0)) %>% # replace NA with 0 (no fire ignitions) 
  # insert `id` column 
  mutate(id = 1:nrow(.),
         .before = fire_count) %>% 
  mutate(fire_prop = fire_count / sum(fire_count)) %>% # no. of fire ignitions in a cell / total no. of fire ignitions
  as.matrix() # turn intro matrix (each column = 1 layer in `RasterBrick`)

# --- vic_raster_crop; with assigned values 
vic_raster_crop_values <- raster::setValues(x = vic_raster_crop,
                                            values = ignition_raster_values)

vic_raster_crop_values_training <- raster::setValues(x = vic_raster_crop,
                                                     values = ignition_raster_values_training)

# --- plot spatial grid results
vic_raster_crop_values %>% 
  plot()

vic_raster_crop_values_training %>% 
  plot()
```

> `training` (historical data) & `predict_x` (satellite data); very different in terms of where fire ignited 

```{r}
# --- side-by-side dot plot
vic_raster_crop_values_df <- vic_raster_crop_values %>% 
  raster::getValues() %>% 
  as_tibble() 

# `fire_count`
vic_raster_crop_values_df %>% 
  ggplot() +
  geom_dotplot(aes(x = factor(id),
                   y = fire_count),
               binaxis = "y",
               stackdir = "center",
               position = "dodge",
               dotsize = 0.3) +
  geom_dotplot(aes(x = factor(id),
                   y = fire_count),
               binaxis = "y",
               stackdir = "center",
               position = "dodge",
               dotsize = 0.3,
               fill = "red",
               data = filter(vic_raster_crop_values_df, fire_count != 0)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   size = 8)) +
  labs(x = "Cell ID",
       y = "Fire Count")

# `fire_prop`
vic_raster_crop_values_df %>% 
  ggplot() +
  geom_dotplot(aes(x = factor(id),
                   y = fire_prop),
               binaxis = "y",
               stackdir = "center",
               position = "dodge",
               dotsize = 0.3) +
  geom_dotplot(aes(x = factor(id),
                   y = fire_prop),
               binaxis = "y",
               stackdir = "center",
               position = "dodge",
               dotsize = 0.3,
               fill = "red",
               data = filter(vic_raster_crop_values_df, fire_prop != 0)) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90,
                                   size = 8)) +
  labs(x = "Cell ID",
       y = "Fire proportion")
```


# Final map (in app) [WIP]
```{r}
# ========== FINAL MAP (in app) *to be made interactive ==========

# === for `predict_x` *satellite data
# --- vic_raster (where risk prediction lies)
final_map <- tmap::tm_shape(vic_raster_crop_values) +
  tmap::tm_raster(title = "bushfire risk predictions",
                  style = "cont",
                  palette = "YlOrRd",
                  n = 5, # no. of colours 
                  alpha = 0.4) + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- draw prediction points (to buffer as necessary)
  tmap::tm_shape(cause_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.5) +
  
  # --- themes
  tmap::tm_layout(legend.outside = T)

tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
final_map
```


```{r}
# ========== FINAL MAP (in app) *to be made interactive ==========

# === for `training` *historical data
# --- vic_raster (where risk prediction lies)

final_map <- tmap::tm_shape(vic_raster_crop_values_training) +
  tmap::tm_raster(title = "bushfire risk predictions",
                  style = "cont",
                  palette = "YlOrRd",
                  n = 5, # no. of colours 
                  alpha = 0.4) + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- draw prediction points (to buffer as necessary)
  tmap::tm_shape(cause_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.5) +
  
  # --- themes
  tmap::tm_layout(legend.outside = T)

tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
final_map
```


# simulation [WIP]
```{r}
# --- get single rectangle resolution
raster::res(vic_raster)

# --- create window; for a grid cell
window <- spatstat.geom::owin(xrange = c(0, 0.45),
                              yrange = c(0, 0.26))

# plot
window %>% 
  plot()

# --- e.g. simulate bushfires; in a grid cell
spatstat.core::rThomas(kappa = 50, # intensity of Poisson process of cluster centers = expected events; per unit area (no. of fires over the historical years / no. of years) *only 2019-2020 reliable spatially
                       # * take hours into a/c
                       
                       scale = 1, # sd of random displacement (along each coordinate axis) of a pt. from its cluster centre
                       
                       mu = 3, # mean no. of points per cluster
                       
                       window = window) %>% # window (unit grid square)
  plot()
```

# EDA plots

## ppp object [WIP]
```{r}
# --- retreive coordinates of predicted bushfires; in matrix form
coords_predict <- cause_sf %>% 
  sf::st_coordinates()

# window
window <- spatstat.geom::owin(xrange = c(140.96168, 149.97629),
        yrange = c(-39.13396, -33.99605))

# --- create ppp object; of predicted pts. 
predict_ppp <- spatstat.geom::ppp(x = coords_predict[,1],
                                  y = coords_predict[,2],
                                  window = window)


# --- plot density
plot(density(predict_ppp))

# --- plot points
plot(predict_ppp)
```






