---
title: "eda"
author: "Brenwin"
date: "06/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# --- load libraries
library(tidyverse)
library(ozmaps)
library(tmap)
library(raster)
library(sf)
library(sp)
library(spatstat)
```

# read in data
```{r, eval=FALSE}
# === read in data sets
load("save.RData")
```

## historical data (`training`) *historical bushfire (less reliable)
```{r}
training <- read_csv("data/training.csv")

training <- training %>%
  filter(!CAUSE %in% c("BURNING BUILDING",
                       "WASTE DISPOSAL, INDUSTRIAL, SAWMILL, TIP",
                       "WASTE DISPOSAL, DOMESTIC",
                       "BURNING VEHICLE, MACHINE",
                       "BURNING BUILDING")) %>%
  filter(new_cause != "other") %>%
  filter(new_cause != "relight")


training <- dplyr::select(training, -c(EVENTID:FIRE_NUM), -id, -CAUSE, -FOREST, -FOR_CODE, -FOR_CAT)

training <- training %>% 
  mutate(year = lubridate::year(FIRE_START),
         month = lubridate::month(FIRE_START),
         day = factor(lubridate::day(FIRE_START), 
                      levels = c(1:31)),
         wod = factor(lubridate::wday(FIRE_START), 
                      levels = c(1:7)))

# training <- training %>% 
#   filter(month %in% c(10,11,12,1,2,3))

training <- na.omit(training)

training <- training %>% 
  mutate(new_cause = ifelse(test = new_cause == "accidental_human",
                            yes = "accident",
                            no = new_cause)) %>%
  mutate(new_cause = ifelse(test = new_cause == "burning_off_human", 
                            yes = "burning_off", 
                            no = new_cause)) %>%
  mutate(new_cause = factor(new_cause)) %>%
  mutate(FOR_TYPE = factor(FOR_TYPE))

training <- na.omit(training)

training <- training %>% 
  mutate(log_dist_cfa = log(dist_cfa),
         log_dist_camp = log(dist_camp),
         log_dist_road = log(dist_road),
         COVER = factor(COVER),
         HEIGHT = factor(HEIGHT))

training <- training %>% 
  rename(cause = new_cause)

training <- mutate(training,
                   cause = forcats::fct_relevel(cause,
                                                "lightning",
                                                "accident",
                                                "arson",
                                                "burning_off"))

training <- na.omit(training)

training <- training %>% 
  dplyr::select(-dist_road, -dist_cfa, -dist_camp, -FIRE_START)
```

## import random forest model (`final_model`)
```{r}
final_model <- readr::read_rds("data/Final_model.rds")
```

## `predict_x` training data *satellite data
```{r}
# Read in predict data
predict_x <- readr::read_csv("data/predict_x.csv")

predict_x <- predict_x %>% 
  mutate(log_dist_cfa = log(dist_cfa),
         log_dist_camp = log(dist_camp),
         log_dist_road = log(dist_road))
```

## use RF model; predict bushfire ignitions cause; in satellite data (`predict_x`) and historical data (`training`)
```{r}
# === use RF model; predict bushfire `cause`

# --- predict cause; in training *historical data 
training_pred <- training %>% 
  # use RF model; predict bushfire cause 
  mutate(cause = predict(final_model,
                         newdata = training)) 

# --- predict cause; in `predict_x` *satellite data
predict_x <- predict_x %>% 
  # use RF model; predict bushfire cause 
  mutate(cause = predict(final_model,
                         newdata = predict_x),
         # incldue `month` and `year` variable
         month = lubridate::month(time),
         year = lubridate::year(time))
```


## join historical & satellite data
```{r}
# --- select relevant variables

# satellite data 2019-2020
cause_df <- predict_x %>% 
  dplyr::select(lon, lat, cause, year, month, FOR_TYPE) %>% 
  mutate(data = "satellite") 

# historical data 
cause_training_df <- training_pred %>% 
  dplyr::select(lon, lat, cause, year, month, FOR_TYPE) %>% 
  mutate(data = "historical")
```


```{r}
cause_join_df <- full_join(cause_df, cause_training_df)

cause_join_df <- cause_join_df %>% 
  mutate(season = case_when(
    month %in% c(12, 1, 2) ~ "summer",
    month %in% c(3, 4, 5) ~ "autumn",
    month %in% c(6, 7, 8) ~ "winter",
    month %in% c(9, 10, 11) ~ "spring"
  )) %>% 
  # mutate(year = factor(year,
  #                      levels = 2000:2020),
  #        month = factor(month,
  #                       levels = 1:12)) %>% 
  arrange(year, month)
```

```{r}
# --- create storage vector
bf_season <- numeric(nrow(cause_join_df))
group <- 1 # start from group 1
current_year <- min(as.numeric(as.character(cause_join_df$year))) # start from minimum year

# --- for loop
# •if following year & month == 10; group + 1
# -> i.e. group by bushfire season (10(Oct) to 3(March))
for(i in seq_along(cause_join_df$year)){
    
  # group + 1; if its next year & month == 10 (October)
  if(current_year == cause_join_df$year[i] - 1 & cause_join_df$month[i]  == 10){
    group <- group + 1
    current_year <- current_year + 1
  }  
  
  bf_season[i] <- group
}

cause_join_df <- cbind(cause_join_df, bf_season) %>% 
  group_by(bf_season) %>% 
  # add bushfire season column
  mutate(bf_season = paste0(min(year), "-", min(year + 1))) %>% 
  ungroup()


cause_join_df <- cause_join_df %>% 
  # not in bushfire season; NA 
  mutate(bf_season = case_when(
    !(month %in% c(1, 2, 3, 10, 11, 12)) ~ NA_character_,
    month %in% c(1, 2, 3, 10, 11, 12) ~ bf_season
  )) 

save(cause_join_df, file = "data/cause_join_df")
```

```{r}
cause_df <- cause_join_df %>% 
  filter(data == "satellite")

cause_training_df <- cause_join_df %>% 
  filter(data == "historical")
```


```{r}
# e.g. === plot results
au_map <- rnaturalearth::ne_states(country = 'Australia', 
                                   returnclass = 'sf')

vic_map <- au_map[7, ]
```

## read in clustering data
```{r readin_clustering}
# --- read in clustering 2016-2021 clustering data
cluster_16_21_df <- readr::read_csv("data/clustering/predict_x_2016_2018.csv") %>% # 2016-2019
  rbind(readr::read_csv("data/clustering/predict_x_2019_2020.csv")) %>% # 2019-2021
  rename(date = time) %>%
  # filter to bushfire season months
  mutate(year = lubridate::year(date),
         month = lubridate::month(date)) %>% 
  filter(month %in% c(1, 2, 3, 10, 11, 12)) %>% 
  arrange(date, year, month)

# ===== add `bf_season` variable (group years from Oct-Mar together)

# --- create storage vector
bf_season <- numeric(nrow(cluster_16_21_df))
group <- 1 # start from group 1
current_year <- min(cluster_16_21_df$year) # start from minimum year

# --- for loop
# •if following year & month == 10; group + 1
# -> i.e. group by bushfire season (10(Oct) to 3(March))
for(i in seq_along(cluster_16_21_df$date)){
    
  # group + 1; if its next year & month == 10 (October)
  if(current_year == cluster_16_21_df$year[i] - 1 & cluster_16_21_df$month[i]  == 10){
    group <- group + 1
    current_year <- current_year + 1
  }  
  
  bf_season[i] <- group
}

cluster_16_21_df <- cbind(cluster_16_21_df, bf_season) %>% 
  group_by(bf_season) %>% 
  # add bushfire season column
  mutate(bf_season = paste0(min(year), "-", min(year + 1))) %>% 
  mutate(bf_season = factor(bf_season)) %>% 
  ungroup()
```

```{r}
# --- convert to sf object 
cluster_16_21_sf <- cluster_16_21_df %>% 
  sf::st_as_sf(coords = c("lon", "lat"))
```


# ignition grid 

Goal
> form spatially weighted grid cells; with each grid cell
  - given ID 
  - based on historical data & 2019-2020 satellite data; compute 
    - fire ignition counts
    - proportion $\frac{\# fire~ ignitions}{\# total}$

> examine the grid cells
  - e.g. is historical data understating bushfire in remote areas?
  - make side-by-side dot plots of proportions; by cell id
  

## create Victoria map `sf` (grid's outline)

> Victoria map; forms spatial grid's outline 

```{r}
# === Victoria map (sfdf MULTIPOLYGON)
vic_map_sf <- ozmaps::ozmap_states %>% 
  filter(NAME == "Victoria")

# --- project crs
vic_map_sf <- sf::st_transform(vic_map_sf,
                               crs = 4326)
```

```{r}
# ====== create `sf` objects

# --- create `cause_sf` ; from predictions *satellite data
cause_sf <- cause_df %>% 
  dplyr::select(lon, lat, cause, year, month, FOR_TYPE, bf_season) %>%  # extract year; from date
  sf::st_as_sf(coords = c("lon", "lat"),
               crs = 4326)

# --- create cause_training_sf (sfdf POINT); *historical (`training`) data
cause_training_sf <- cause_training_df %>% 
  dplyr::select(lon, lat, cause, year, month, FOR_TYPE, bf_season) %>% 
  sf::st_as_sf(coords = c("lon", "lat"),
               crs = 4326) %>% 
  filter(month %in% c(1, 2, 3, 10, 11, 12))

# --- joint data
cause_join_sf <- cause_join_df %>% 
  sf::st_as_sf(coords = c("lon", "lat"))
```

## create spatial grid `RasterBrick`


### create `RasterBrick`; cropped to Victoria map window

> `RasterBrick`; designed; store multiple raster (with same extents(bbox) & dimensions)
  - i.e. multi-layer raster

```{r}
# === vic_raster 

# --- create `vic_raster` object *to be toggled
vic_raster <- raster::brick(
  # no. of rows & columns (directly linked to resolution of grid cell)
  nrows = 20,
  ncols = 20,
  
  # bbox (bounding box of Victoria)
  xmn = 140.9617,
  xmx = 149.9763,
  ymn = -39.13396,
  ymx = -33.99605,
  
  # crs
  crs = 4326,
  
  # set `raster` values (rowwise)
  # vals = seq(from = 1, to = 400000, by = 1000) 
  )

# --- mask raster; to only Victorian map 

# change vic_map_sf to `sp` object 
# *`raster` package; NOT compatible with `sf` yet; so; need; change to `sp` 
vic_map_sp <- as(vic_map_sf, 
                 Class = "Spatial")

# mask (*think: crop to polygon shape) raster; to only Victorian map (`vic_map_sp`)
vic_raster_crop <- vic_raster %>% 
  raster::mask(mask = vic_map_sp)
```

### ignition Rules [WIP]
> to be set; when more data and information on appropriate rules are determined
   - e.g. filter to seasons 
    - fire; more prone in summer (increase bushfire risk)
   - e.g. categorise each grid cell; according to fuel(vegetation)
    - e.g. fire cannot occur on water (no bushfire risk)

```{r}
ignition_rules_raster <- raster::brick(
  # no. of rows & columns (directly linked to resolution of grid cell)
  nrows = 20,
  ncols = 20,
  
  # bbox (bounding box of Victoria)
  xmn = 140.9617,
  xmx = 149.9763,
  ymn = -39.13396,
  ymx = -33.99605,
  
  # crs
  crs = 4326,
  
  # set `raster` values (rowwise)
  # vals = seq(from = 1, to = 400000, by = 1000) 
  )

# --- convert `raster` => `spPolygonsdf` => `sf` Polygons`; to conduct spatial join 
ignition_rules_sp <- as(ignition_rules_raster, "SpatialPolygonsDataFrame")
ignition_rules_sf <- sf::st_as_sf(ignition_rules_sp) 

ignition_rules_sf_values <- sf::st_join(ignition_rules_sf, cause_sf) 
```

```{r}
# --- plot ignition rule 
tmap::tm_shape(ignition_rules_sf_values) +
  tm_polygons(col = "FOR_TYPE") +
  tm_shape(vic_map_sf) +
  tm_borders(lwd = 3)
```


### Initial ignition grid: Set values to each grid cell

> set values to ignition grid cell
  - `fire_count`: count the number of fire ignitions; per grid cell
  - `fire_prop`: # ignitions in grid cell / # total no. of ignitions

* note: good to separate satellite data (`predict_x`) with historical data (`training_pred`)
  - so; able; parse them out; when examining grid 

```{r}
# ========== convert `sf` -> `spdf` -> SpatialPoints; to be able; use `rasterize` (count; no. of pts.; in each raster cell)

# ---------- predict_x *satellite data
cause_spdf <- as(cause_sf, Class = "Spatial")
cause_sp <- as(cause_spdf, "SpatialPoints")

# ---------- training_pred *historical data

# will filter `sf` object; before turning int `spdf` object in app  
# e.g. filter to year 2000 only (user-specification!)
cause_training_sf1 <- cause_training_sf %>% 
  filter(year == 2018,
         month %in% c(1, 2, 3, 10, 11, 12)) # same months as satellite data 

cause_training_spdf <- as(cause_training_sf1, Class = "Spatial")
cause_training_sp <- as(cause_training_spdf, "SpatialPoints")

# ---------- join data 
cause_join_spdf <- as(cause_join_sf, Class = "Spatial")
cause_join_sp <- as(cause_join_spdf, "SpatialPoints")
```

> NOTE: BEFORE creating `spdf` object; is where; filter to `year` etc. (user specifications)

```{r}
# === for satellite & historical data *join

# --- count the no. of points in each `raster` `vic_raster_crop` cell
ignition_join_rasterize <- raster::rasterize(x = cause_join_sp,
                                             y = vic_raster_crop,
                                             fun = "count",
                                             background = 0) # set cells w. no fire ignitions to 0

# ----- compute; in each grid cell
# • `id`: id for cell (rowwise)
# • `fire_count`: no. of points
# • `fire_prop`: no. of points / total points (bushfire ignitions)
# *returns matrix

ignition_raster_values_join <- raster::getValues(ignition_join_rasterize) %>% 
  tibble(fire_count = .) %>% 
  mutate(id = 1:nrow(.), # cell id 
         .before = fire_count) %>% 
  mutate(fire_prop = fire_count / sum(fire_count)) %>% # no. of fire ignitions in a cell / total no. of fire ignitions
  as.matrix() # turn intro matrix (each column = 1 layer in `RasterBrick`)

vic_raster_crop_values_join <- raster::setValues(x = vic_raster_crop,
                                                 values = ignition_raster_values_join)

vic_raster_crop_values_join %>% 
  plot()
```


```{r}
# --- count the number of points in each raster cell; via `raster::rasterize`

# predict_x *satellite data 
ignition_rasterize <- raster::rasterize(x = cause_sp,# `SpatialPoints` object
                                        y = vic_raster_crop, # `Raster` object 
                                        fun = "count",
                                        background = 0) # set cells w. no fire ignitions to 0


ignition_rasterize_training <- raster::rasterize(x = cause_training_sp,# `SpatialPoints` object
                                                 y = vic_raster_crop, # `Raster` object 
                                                 fun = "count",
                                                 background = 0) # set cells w. no fire ignitions to 0

# ----- compute; in each grid cell
# • `id`: id for cell (rowwise
# • `fire_count`: no. of points
# • `fire_prop`: no. of points / total points (bushfire ignitions)

# predict_x *satellite data 
ignition_raster_values <- raster::getValues(ignition_rasterize) %>% # extract count values
  tibble(fire_count = .) %>% 
  mutate(id = 1:nrow(.),
         .before = fire_count) %>% 
  mutate(fire_prop = fire_count / sum(fire_count)) %>% # no. of fire ignitions in a cell / total no. of fire ignitions
  as.matrix() # turn intro matrix (each column = 1 layer in `RasterBrick`)

# `training` *historical data
ignition_raster_values_training <- raster::getValues(ignition_rasterize_training) %>% # extract count values
  tibble(fire_count = .) %>% 
  # insert `id` column 
  mutate(id = 1:nrow(.),
         .before = fire_count) %>% 
  mutate(fire_prop = fire_count / sum(fire_count)) %>% # no. of fire ignitions in a cell / total no. of fire ignitions
  as.matrix() # turn intro matrix (each column = 1 layer in `RasterBrick`)

# --- vic_raster_crop; with assigned values 
vic_raster_crop_values <- raster::setValues(x = vic_raster_crop,
                                            values = ignition_raster_values)

vic_raster_crop_values_training <- raster::setValues(x = vic_raster_crop,
                                                     values = ignition_raster_values_training)

# --- plot spatial grid results
vic_raster_crop_values %>% 
  plot()

vic_raster_crop_values_training %>% 
  plot()
```

> `training` (historical data) & `predict_x` (satellite data); very different in terms of where fire ignited 

# EDA plots

## Frequency distribution for number of fires ignited

> for each simulation; draw the no. of bushfires ignited per time period
  - time period can be based on
    - seasons (summer, autumn, winter, spring)
    - year 
    
```{r}
# --- frequency distribution; based on year & season
freq_dist_year_season <- cause_join_df %>% 
  filter(season %in% c("summer")) %>% 
  count(year)
```
    

### EDA plots: `cause` by `year` bar plot (temporal)
```{r}
# ---------- `predict_x` *satellite data

# --- `cause` by `year`
cause_join_df %>% 
  filter(month %in% c(1,2,3,10,11,12)) %>% 
  group_by(year) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(year,
                          levels = c(2000:2020)),
               y = total)) +
  geom_col(aes(x = factor(year),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))


# --- `cause` by `month`
predict_x %>% 
  filter(month %in% c(1,2,3,10,11,12)) %>%
  group_by(month) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(month),
               y = total)) +
  geom_col(aes(x = factor(month),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw()

# --- `cause` by bf_season
cause_join_df %>% 
  filter(!is.na(bf_season)) %>% 
  group_by(bf_season) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(bf_season),
               y = total)) +
  geom_col(aes(x = factor(bf_season),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```


## Simulation

### Simulate ACROSS cells *for bushfire ignitions
```{r}
# ========== e.g. 1,000 simulation ==========
# --- no. of fires; generated from uniform dist. 

# --- create empty storage matrix 
sim_count <- matrix(NA, 
                    nrow = 0,
                    ncol = 3)

# --- run for loop: simulate random points (bushfire ignitions); in `raster` cell; return coordinates (matrix)
for(i in 1:1000){
  # simulate random points; in `raster` cells  
  sim <- enmSdm::sampleRast(x = vic_raster_crop_values$fire_count, # `raster` object
                            # *** draw no. of points; from frequency distribution *** 
                            n = sample(freq_dist_year_season$n, 
                                       size = 1), # 1 number
                            replace = T, # w/o replacement
                            prob = T) # sample cells with probabilities proportional to cell value (fire_count)
  
  # include 3rd column: simulation iteration number
  sim <- cbind(sim, rep(i, nrow(sim)))
  
  # store results in `sim_count` 
  sim_count <- rbind(sim_count, sim)
}

# create `SpatialPoints` object to store simulation points; 
# so; able; `rasterize` (count no. of points in each cell)
sim_count_sp <- sim_count %>% 
  as_tibble() %>% 
  rename(lon = x,
         lat = y,
         sim_id = V3) %>%
  select(-sim_id) %>% # cannot have `sim_id` when `rasterize`
  # convert to `sp` object
  sp::SpatialPoints()

# --- count number of points(bushfire ignitions); in each cell 
sim_count_raster <- raster::rasterize(x = sim_count_sp,# `SpatialPoints` object
                                      y = vic_raster_crop, # `Raster` object 
                                      fun = "count") 

# === plot simulation results

# --- using `plot`
sim_count_raster %>% 
  plot()

# --- using `tmap` 

#  convert `raster` -> `spdf` -> `sf` object for plotting
sim_count_raster_sf <- as(sim_count_raster, "SpatialPolygonsDataFrame") %>% 
  sf::st_as_sf() %>% 
  rename(fire_count = layer) %>% 
  mutate(fire_count = fire_count / 1000) # divide `fire_count`; by no. of simulations
  
tmap::ttm() 

tmap::tm_basemap(leaflet::providers$Esri.WorldTopoMap) +
tmap::tm_shape(sim_count_raster_sf) + 
  tmap::tm_polygons(col = "fire_count") +
  # include outline of victoria map 
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 2)
```
> to fill in: frequency distribution; from which; number of bushfires occur 

### Simulate WITHIN cell *potentially for bushfire spreads [WIP]

> `ppp` objects; useful; simulate WITHIN cell

```{r}
# --- retreive coordinates of predicted bushfires; in matrix form
coords_predict <- cause_sf %>% 
  sf::st_coordinates()

# window
window <- spatstat.geom::owin(xrange = c(140.96168, 149.97629),
        yrange = c(-39.13396, -33.99605))

# --- create ppp object; of predicted pts. 
predict_ppp <- spatstat.geom::ppp(x = coords_predict[,1],
                                  y = coords_predict[,2],
                                  window = window)

# --- plot density
plot(density(predict_ppp))

# --- plot points
plot(predict_ppp)
```

```{r}
# --- get single rectangle resolution
raster::res(vic_raster)

# --- create window; for a grid cell
window <- spatstat.geom::owin(xrange = c(0, 0.45),
                              yrange = c(0, 0.26))

# plot
window %>% 
  plot()

# --- e.g. simulate bushfires; in a grid cell
spatstat.core::rThomas(kappa = 50, # intensity of Poisson process of cluster centers = expected events; per unit area (no. of fires over the historical years / no. of years) *only 2019-2020 reliable spatially
                       # * take hours into a/c
                       
                       scale = 1, # sd of random displacement (along each coordinate axis) of a pt. from its cluster centre
                       
                       mu = 3, # mean no. of points per cluster
                       
                       window = window) %>% # window (unit grid square)
  plot()
```

ONE SIMULATION
1. Simulate number of fires from a frequency distribution (based on historical no. of fires data)
-> e.g. 5 fires

2. simulate ignition locations; based on spatially probability weighted weighted grid cells 
-> incorporate ignition rules 
-> historical fire data

*potentially simulate burning conditions for each individual simulated fire 
-> based on weather variables// simulate the no. of days; fire burn (based on frequency distribution)
=> gauge the fire spread 

bushfire risk = no. of times cell burned / no. of simulations


## deal with discrepancy between historical & satellite data

> historical data; can be thought as: where 1st responder saw the fire 
> hotspot data; detects ignition using satellite; able; detect ignitions in remote areas
  - ∴ result in discrepancies; esp. in remote areas

> goal: find a statistic; appropriately deal with the discrepency 

### satellite (`predict_x`) against historical data (`training`)

> compare agreeableness of 2018 historical data vs. 2019-2020 satellite hotspot data

```{r}
ignition_raster_values_df <- as_tibble(ignition_raster_values) %>% 
  # select(-id) %>% 
  mutate(data = "sattelite")

ignition_raster_values_training_df <- as_tibble(ignition_raster_values_training) %>% 
  # select(-id) %>% 
  mutate(data = "historical")


# --- plot `fire_count` of historical data against statellite data
# examine agreeablenesss 
ggplot() +
  geom_text(aes(x = ignition_raster_values_df$fire_count,
                y = ignition_raster_values_training_df$fire_count,
                label = factor(1:400)), # id 1 to 400 
            alpha = 0.6) +
  theme_bw() +
  labs(x = "satellite data",
       y = "historical data")
```

```{r}
ignition_raster_values_join <- as_tibble(ignition_raster_values_join) %>% 
  # include historical - satellite fire ignitions in a cell
  mutate(fire_count_diff = ignition_raster_values_training_df$fire_count - ignition_raster_values_df$fire_count) %>% # 
  as.matrix() # turn intro matrix (each column = 1 layer in `RasterBrick`)

# --- set values; to raster
vic_raster_crop_values_join <- raster::setValues(x = vic_raster_crop,
                                                 values = ignition_raster_values_join)

# --- convert `raster` -> `spdf` -> `sf` for plotting with `tmap`
# *sf polygons; allow popup of variable upon hover
# `tm_raster` does not allow
vic_crop_values_join_sp <- as(vic_raster_crop_values_join, "SpatialPolygonsDataFrame")
vic_crop_values_join_sf <- sf::st_as_sf(vic_crop_values_join_sp) 

# --- plot `sf` object; 
p <- tmap::tm_shape(vic_crop_values_join_sf) +
  tm_polygons(col = "fire_count_diff", # fill by historical - satellite fire ignitions; in grid cell
              palette = "RdBu",
              n = 7,
              group = "fire_count_diff",
              id = "id",
              alpha = 0.3) +
  # include Victoria map outline
  tm_shape(vic_map_sf) +
  tm_borders(lwd = 3)

# tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
p
```

> blue: historical > satellite 
> red: satellite > historical

### historical data dot plots  

#### by year
```{r}
# --- create function; count; no. of points in each raster cell; via raster::rasterize; for each year 
extract_ignition_raster_values_year <- function(data, year){
  
  # --- count the number of points in each raster cell; via `raster::rasterize`  
  
  cause_sf <- data %>% 
    filter(year == {{year}},
           month %in% c(1, 2, 3, 10, 11, 12)) # same months as satellite data *bushfire season
  
  # convert from sf -> spdf -> sp; to use `raster::rasterize`
  cause_spdf <- as(cause_sf, Class = "Spatial")
  cause_sp <- as(cause_spdf, "SpatialPoints")
  
  # count; no. of points; in each raster cell
  ignition_rasterize <- raster::rasterize(x = cause_sp,# `SpatialPoints` object
                                          y = vic_raster_crop, # `Raster` object 
                                          fun = "count",
                                          background = 0) # set cells w. no fire ignitions to 0
  
  # ----- compute; in each grid cell
  # • `id`: id for cell (rowwise
  # • `fire_count`: no. of points
  
  # --- extract grid cells values; into each tibble
  ignition_raster_values <- raster::getValues(ignition_rasterize) %>% # extract count values
    tibble(fire_count = .) %>% 
    # insert `id` column 
    mutate(id = 1:nrow(.),
           year = year,
           .before = fire_count) 
  
  return(ignition_raster_values)
}

ignition_raster_values_training_years <- purrr::map_df(.x = 2000:2018,
                                                       .f = ~extract_ignition_raster_values_year(cause_training_sf, .x)) 


# plot
p1 <- ignition_raster_values_training_years %>% 
  ggplot() +
  geom_point(aes(x = factor(year),
                 y = fire_count)) +
  facet_wrap(~ id,
             nrow = 20,
             ncol = 20) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
p1
# plotly::ggplotly(p1)
```

```{r}
# --- filter to cells(`id`); where historical(2018) > satellite(2019-2020); by 20
hist_gt_sat <- as_tibble(ignition_raster_values_join) %>% 
  # include historical - satellite fire ignitions in a cell
  mutate(fire_count_diff = ignition_raster_values_training_df$fire_count - ignition_raster_values_df$fire_count) %>% 
  filter(fire_count_diff > 20) %>% 
  pull(id)

ignition_raster_values_training_years %>% 
  filter(id %in% hist_gt_sat) %>% 
  ggplot() +
  geom_point(aes(x = factor(year),
                 y = fire_count)) +
  facet_wrap(~ id) +
  theme_bw()
  
# --- filter to cells(`id`); where satellite(2019-2020) > historical(2018); by 20
sat_gt_hist <- as_tibble(ignition_raster_values_join) %>% 
  # include historical - satellite fire ignitions in a cell
  mutate(fire_count_diff = ignition_raster_values_training_df$fire_count - ignition_raster_values_df$fire_count) %>% 
  filter(fire_count_diff < -15) %>% 
  pull(id)

ignition_raster_values_training_years %>% 
  filter(id %in% sat_gt_hist) %>% 
  ggplot() +
  geom_point(aes(x = factor(year),
                 y = fire_count)) +
  facet_wrap(~ id) +
  theme_bw()
```

> see in conjunection with map above! 

> many cells; do not have fires historically 

> for cells; with historical ignitions > satellite ignitions
  - there seem to be a spike in historical data ignitions; from 2014 onwards; to 2018; as compared to previous years
  - 2018 showed a significant increase in fire ignitions compared to all other years; for all cells

> cell 208: outlier 
  - has more ignitions; in early years
  .˙. remove from counts

> ???could it be; data; more consistently recorded; in later years?
  
> for cells with satellite ignitions > historical ignitions
  - in the South East; there were still counts in historical data; although minimal (fluctuate from 5-10 mostly) *close to roads (supposedly; caused by accidents rather than lightning)
  - in SouthWest; interesting that very few fires were ignited; in historical data; but many captured on satellite data 

> seems like; smth. to do with watchtowers 

#### by bushfire season
```{r}
# --- create funciton; count; no. of points in each raster cell; via raster::rasterize; for each year 
extract_ignition_raster_values_bf_season <- function(data, bf_season){
  
  # --- count the number of points in each raster cell; via `raster::rasterize`  
  
  cause_sf <- data %>% 
    filter(bf_season == {{bf_season}}) # same months as satellite data *bushfire season
  
  # convert from sf -> spdf -> sp; to use `raster::rasterize`
  cause_spdf <- as(cause_sf, Class = "Spatial")
  cause_sp <- as(cause_spdf, "SpatialPoints")
  
  # count; no. of points; in each raster cell
  ignition_rasterize <- raster::rasterize(x = cause_sp,# `SpatialPoints` object
                                          y = vic_raster_crop, # `Raster` object 
                                          fun = "count",
                                          background = 0) # set cells w. no fire ignitions to 0
  
  # ----- compute; in each grid cell
  # • `id`: id for cell (rowwise
  # • `fire_count`: no. of points
  
  # --- extract grid cells values; into each tibble
  ignition_raster_values <- raster::getValues(ignition_rasterize) %>% # extract count values
    tibble(fire_count = .) %>% 
    # insert `id` column 
    mutate(id = 1:nrow(.),
           bf_season = bf_season,
           .before = fire_count) 
  
  return(ignition_raster_values)
}
```

```{r}
# ============ For training data ============
ignition_raster_values_training_bf_season <- purrr::map_df(.x = unique(cause_training_sf$bf_season),
                                                       .f = ~extract_ignition_raster_values_bf_season(cause_training_sf, .x)) 

# plot
p1 <- ignition_raster_values_training_bf_season %>% 
  ggplot() +
  geom_point(aes(x = factor(bf_season),
                 y = fire_count)) +
  facet_wrap(~ id,
             nrow = 20,
             ncol = 20) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
p1
# plotly::ggplotly(p1)
```

```{r}
# ============ for satellite data; 2016 - 2021 ============
ignition_raster_values_cluster_bf_season <- purrr::map_df(.x = unique(cluster_16_21_sf$bf_season),
                                                       .f = ~extract_ignition_raster_values_bf_season(cluster_16_21_sf, .x)) 

p1 <- ignition_raster_values_cluster_bf_season %>% 
  ggplot() +
  geom_point(aes(x = factor(bf_season),
                 y = fire_count)) +
  facet_wrap(~ id,
             nrow = 20,
             ncol = 20) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
p1
```

```{r}
# --- join both data
ignition_raster_values_join_bf_season <- inner_join(ignition_raster_values_cluster_bf_season, ignition_raster_values_training_bf_season,
           by = c("bf_season", "id")) %>% 
  rename(fire_count_sat = fire_count.x,
         fire_count_hist = fire_count.y)

p <-ignition_raster_values_join_bf_season %>% 
  ggplot() +
  geom_text(aes(x = fire_count_sat,
                y = fire_count_hist,
                label = id)) +
  facet_wrap(~ bf_season,
             nrow = 3) +
  theme_bw() 
  
plotly::ggplotly(p)
```


### comparing maps of historical vs. satellite data *spatial

#### by year
```{r}
# --- function; extract `sf` polygon; for each grid cell; to facet by `year`

extract_ignition_raster_values_sf_year <- function(data, year){ # data: sf object
  
  # --- count the number of points in each raster cell; via `raster::rasterize`  
  cause_sf <- data %>% 
    filter(year == {{year}},
           month %in% c(1, 2, 3, 10, 11, 12)) # same months as satellite data *bushfire season
  
  
  cause_spdf <- as(cause_sf, Class = "Spatial")
  cause_sp <- as(cause_spdf, "SpatialPoints")
  
  ignition_rasterize <- raster::rasterize(x = cause_sp,# `SpatialPoints` object
                                          y = vic_raster_crop, # `Raster` object 
                                          fun = "count",
                                          background = 0) # set cells w. no fire ignitions to 0
  
  # ----- compute; in each grid cell
  # • `fire_count`: no. of points
  # • `fire_prop`: no. of points / total points (bushfire ignitions)
  
  # raster -> spdf -> sf polygons 
   ignition_rasterize_sp <- as(ignition_rasterize, "SpatialPolygonsDataFrame")
   
   ignition_rasterize_training_sf <- sf::st_as_sf(ignition_rasterize_sp) %>%
     rename(fire_count = layer) %>% 
     mutate(id = 1:400, # add id column; for each raster cel id 
            year = {{year}}) # include `year` column
   
   return(ignition_rasterize_training_sf)
}

# --- extract list of `sf` polygons; for year 2000 to 2018
ignition_rasterize_training_sf_year <- purrr::map(.x = 2000:2018,
                                             .f = ~extract_ignition_raster_values_sf_year(cause_training_sf, .x))

# --- bind list of `sf` polygon objects into 1
ignition_rasterize_training_sf_year <- dplyr::bind_rows(ignition_rasterize_training_sf_year)
```


```{r}
# === for `training` *historical data
# --- vic_raster (where risk prediction lies)
hist_map_facet_year <- tmap::tm_shape(ignition_rasterize_training_sf_year) +
  tmap::tm_polygons(col = "fire_count",
                    id = "id",
                    style = "cont",
                    palette = "YlOrRd",
                    n = 5, # no. of colours 
                    alpha = 0.4,
                    legend.show = F) + 
  tmap::tm_facets(by = "year") + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- add ignition points (to buffer as necessary)
  tmap::tm_shape(cause_training_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.1) +
  tmap::tm_facets(by = "year") +
  # --- themes
  tmap::tm_layout(legend.show = F)


# tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
hist_map_facet_year
```

```{r}
# === for satellite data 

# --- convert `raster` -> `spdf` -> `sf` poly; for plotting
ignition_rasterize_sf <- as(ignition_rasterize, "SpatialPolygonsDataFrame") %>% 
  sf::st_as_sf(.) %>% 
  rename(fire_count = layer) %>% 
  mutate(id = 1:400,
         year = 2019) # 2019-2020 bushfire season

sat_map <- tmap::tm_shape(ignition_rasterize_sf) +
  tmap::tm_polygons(col = "fire_count",
                    id = "id",
                    style = "cont",
                    palette = "YlOrRd",
                    n = 5, # no. of colours 
                    alpha = 0.4,
                    legend.show = T) + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- draw prediction points (to buffer as necessary)
  tmap::tm_shape(cause_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.1) +
  # --- themes
  tmap::tm_layout(legend.show = T)

sat_map
```

```{r}
# --- plot `avg_fire_count` (avg. no. of ignitions; in each cell)
ignition_rasterize_training_sf_avg <- ignition_rasterize_training_sf_year %>% 
  group_by(id) %>% 
  summarise(avg_fire_count = mean(fire_count)) 

hist_map_avg <- tmap::tm_shape(ignition_rasterize_training_sf_avg) +
  tmap::tm_polygons(col = "avg_fire_count",
                    id = "id",
                    style = "cont",
                    palette = "YlOrRd",
                    n = 5, # no. of colours 
                    alpha = 0.4,
                    legend.show = F) + 
  # tmap::tm_facets(by = "year") + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- add ignition points (to buffer as necessary)
  tmap::tm_shape(cause_training_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.1) +
  # tmap::tm_facets(by = "year") +
  # --- themes
  tmap::tm_layout(legend.show = F)

hist_map_avg
```

```{r}
# --- plot  median `median_fire_count` (median no. of ignitions; in each cell)
ignition_rasterize_training_sf_median <- ignition_rasterize_training_sf_year %>% 
  group_by(id) %>% 
  summarise(median_fire_count = median(fire_count)) 

hist_map_median <- tmap::tm_shape(ignition_rasterize_training_sf_median) +
  tmap::tm_polygons(col = "median_fire_count",
                    id = "id",
                    style = "cont",
                    palette = "YlOrRd",
                    n = 5, # no. of colours 
                    alpha = 0.4,
                    legend.show = T) + 
  # tmap::tm_facets(by = "year") + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- add ignition points (to buffer as necessary)
  tmap::tm_shape(cause_training_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.1) +
  # tmap::tm_facets(by = "year") +
  # --- themes
  tmap::tm_layout(legend.show = F)

hist_map_median
```

#### by bushfire season
```{r}
# --- function; extract `sf` polygon; for each grid cell; to facet by `year`

extract_ignition_raster_values_sf_bf_season <- function(data, bf_season){ # data: sf object
  
  # --- count the number of points in each raster cell; via `raster::rasterize`  
  cause_sf <- data %>% 
    filter(bf_season == {{bf_season}}) # same months as satellite data *bushfire season
  
  
  cause_spdf <- as(cause_sf, Class = "Spatial")
  cause_sp <- as(cause_spdf, "SpatialPoints")
  
  ignition_rasterize <- raster::rasterize(x = cause_sp,# `SpatialPoints` object
                                          y = vic_raster_crop, # `Raster` object 
                                          fun = "count",
                                          background = 0) # set cells w. no fire ignitions to 0
  
  # ----- compute; in each grid cell
  # • `fire_count`: no. of points
  # • `fire_prop`: no. of points / total points (bushfire ignitions)
  
  # raster -> spdf -> sf polygons 
   ignition_rasterize_sp <- as(ignition_rasterize, "SpatialPolygonsDataFrame")
   
   ignition_rasterize_training_sf <- sf::st_as_sf(ignition_rasterize_sp) %>%
     rename(fire_count = layer) %>% 
     mutate(id = 1:400, # add id column; for each raster cel id 
            bf_season = {{bf_season}}) # include `year` column
   
   return(ignition_rasterize_training_sf)
}
```

```{r}
# ========== for `training` *historical data ==========

# --- extract list of `sf` polygons; for year 2000 to 2018
ignition_rasterize_training_sf_bf_season <- purrr::map(.x = unique(cause_training_sf$bf_season),
                                                       .f = ~extract_ignition_raster_values_sf_bf_season(cause_training_sf, .x))

# --- bind list of `sf` polygon objects into 1 df
ignition_rasterize_training_sf_bf_season <- dplyr::bind_rows(ignition_rasterize_training_sf_bf_season)

# --- plot results; (where risk prediction lies)
hist_map_facet_bf_season <- tmap::tm_shape(ignition_rasterize_training_sf_bf_season) +
  # --- cells; filled with no. of `fire_count`
  tmap::tm_polygons(col = "fire_count",
                    id = "id",
                    style = "cont",
                    palette = "YlOrRd",
                    n = 5, # no. of colours 
                    alpha = 0.4,
                    legend.show = F) + 
  
  tmap::tm_facets(by = "bf_season",
                  ncol = 3) + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- add ignition points (to buffer as necessary)
  tmap::tm_shape(cause_training_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.1) +
  tmap::tm_facets(by = "bf_season",
                  ncol = 3) +
  # --- themes
  tmap::tm_layout(legend.show = F)


# tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
hist_map_facet_bf_season
```

```{r}
# ========== for satellite data; 2016-2021 ==========

# --- extract list of `sf` polygons; for year 2000 to 2018
ignition_rasterize_cluster_sf_bf_season <- purrr::map(.x = unique(cluster_16_21_sf$bf_season),
                                                       .f = ~extract_ignition_raster_values_sf_bf_season(cluster_16_21_sf, .x))

# --- bind list of `sf` polygon objects into 1 df
ignition_rasterize_cluster_bf_season <- dplyr::bind_rows(ignition_rasterize_cluster_sf_bf_season)


# === plot results 
cluster_map_facet_bf_season <- tmap::tm_shape(ignition_rasterize_cluster_bf_season) +
  tmap::tm_polygons(col = "fire_count",
                    id = "id",
                    style = "cont",
                    palette = "YlOrRd",
                    n = 5, # no. of colours 
                    alpha = 0.4,
                    legend.show = F) + 
  
  tmap::tm_facets(by = "bf_season") + 
  
  # --- add ignition points (to buffer as necessary)
  tmap::tm_shape(cluster_16_21_sf) +
  tmap::tm_bubbles(# col = "cause",
                   size = 0.3,
                   alpha = 0.1,
                   id = "") +
  tmap::tm_facets(by = "bf_season") +
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- themes
  tmap::tm_layout(legend.show = F)

cluster_map_facet_bf_season
```


### comparing historical vs. satellite: plot `cause` by `year` bar plot 

#### by bushfire season
```{r}
# === 2019-2020: satellite data
# === 2000:2019: historical data 

# --- `cause` by `bf_season`
cause_join_count_bf_season <- cause_join_df %>% 
  filter(!is.na(bf_season)) %>% 
  group_by(bf_season) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ungroup() 

cause_join_count_bf_season %>% 
  ggplot() +
  geom_col(aes(x = factor(bf_season),
               y = total)) +
  geom_col(aes(x = factor(bf_season),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "historical data 2000-2019// sat. data 2019-2020",
       x = "bushfire season",
       y = "number of ignitions")

# === cluster data
cluster_16_21_count_bf_season <- cluster_16_21_df %>% 
  filter(!is.na(bf_season)) %>% 
  group_by(bf_season) %>% 
  count() %>% 
  ungroup()  
  # mutate(total = sum(n)) *no `cause`
  
cluster_16_21_count_bf_season %>% 
  ggplot() +
  geom_col(aes(x = factor(bf_season),
               y = n)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "satellite data",
       x = "bushfire season",
       y = "number of ignitions")
```

> the large discrepancy in historical & satellite data; probably due to; significant increase in `accident` ignitions; from 2014 to 2018  

```{r}
cause_join_count_bf_season %>% 
  filter(bf_season %in% unique(cluster_16_21_count_bf_season$bf_season)) %>% 
  select(bf_season, total) %>% 
  # join cluster & join data counts
  right_join(cluster_16_21_count_bf_season,
            by = "bf_season") %>% 
  rename(total_join = total,
         total_cluster = n) %>% # fire ignitions; in clustering data (2016-2021)
  distinct(bf_season,
           .keep_all = T) %>% 
  pivot_longer(cols = total_join:total_cluster,
               names_to = "data", # in join (historical(2000-2019) & sat(2019-2020))// cluster(2016-2021)
               values_to = "n") %>% # no. of fire ignitions
  ggplot() +
  geom_col(aes(x = bf_season,
               y = n,
               fill = data),
           position = "dodge") +
  colorspace::scale_fill_discrete_qualitative() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))
```

> large discrepancy; 2017 - 2018 

```{r}
# --- `cause` by `month`
cause_join_df %>% 
  # filter(bf_season %in% unique(cluster_16_21_df$bf_season))
  group_by(month) %>% 
  count(cause) %>% 
  mutate(total = sum(n)) %>% 
  ggplot() +
  geom_col(aes(x = factor(month),
               y = total)) +
  geom_col(aes(x = factor(month),
               y = n,
               fill = cause)) +
  facet_wrap(. ~ cause) +
  theme_bw()

cause_join_count_month <- cause_join_df %>% 
  filter(bf_season %in% unique(cluster_16_21_df$bf_season)) %>% 
  group_by(bf_season, month) %>% 
  count() %>% 
  ungroup()

cluster_16_21_count_month <- cluster_16_21_df %>% 
  group_by(month, bf_season) %>% 
  count() %>% 
  ungroup()

setdiff(cluster_16_21_count_month$bf_season, cause_join_count_month$bf_season)

left_join(cluster_16_21_count_month, cause_join_count_month,
          by = c("bf_season", "month")) %>% 
  rename(total_cluster = n.x,
         total_join = n.y) %>% 
  pivot_longer(cols = total_cluster:total_join,
               names_to = "data",
               values_to = "n") %>% 
  ggplot() +
  geom_col(aes(x = factor(month,
                          levels = c(10, 11, 12, 1, 2, 3)),
               y = n,
               fill = data),
           position = "dodge") +
  facet_wrap(. ~ bf_season) + 
  colorspace::scale_fill_discrete_qualitative()
```

> missing data; join data 2019: 1, 2, 3
  - i.e. 2019 1, 2, 3 data 


# Final map (in app) [WIP]
```{r}
# ========== FINAL MAP (in app) *to be made interactive ==========

# === for `predict_x` *satellite data
# --- vic_raster (where risk prediction lies)
final_map <- tmap::tm_shape(vic_raster_crop_values_training$fire_count) +
  tmap::tm_raster(title = "bushfire risk predictions",
                  style = "cont",
                  palette = "YlOrRd",
                  n = 5, # no. of colours 
                  alpha = 0.4) + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- draw prediction points (to buffer as necessary)
  tmap::tm_shape(cause_training_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.3) +
  tmap::tm_facets(by = "year") +
  # --- themes
  tmap::tm_layout(legend.outside = T)

tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
final_map
```


```{r, eval=FALSE}
# ========== FINAL MAP (in app) *to be made interactive ==========

# === for `training` *historical data
# --- vic_raster (where risk prediction lies)

final_map <- tmap::tm_shape(vic_raster_crop_values_join) +
  tmap::tm_raster(title = "bushfire risk predictions",
                  style = "cont",
                  palette = "YlOrRd",
                  n = 5, # no. of colours 
                  alpha = 0.4) + 
  
  # --- vic_map_sf (draw outline of Victoria )
  tmap::tm_shape(vic_map_sf) +
  tmap::tm_borders(lwd = 3) + # border line width 
  
  # --- draw prediction points (to buffer as necessary)
  tmap::tm_shape(cause_sf) +
  tmap::tm_bubbles(col = "cause",
                   size = 0.3,
                   alpha = 0.5) +
  
  # --- themes
  tmap::tm_layout(legend.outside = T)

# tmap::ttm() # set tmap mode; to from "plot" (static plot) => to interactive viewing "view"
final_map
```





